Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-03 paper-2024-sac-marl-neighbouring-strategies.tex
update sota with centralized approaches

\sloppy
Many prominent approaches in this field employ a centralized training and decentralized execution paradigm
 (e.g., COMA~\cite{foerster2018counterfactual}, MAPPO~\cite{yu2022surprising}, MADDPG~\cite{DBLP:conf/nips/LoweWTHAM17}, QMIX~\cite{DBLP:conf/icml/RashidSWFFW18}, QTRAN~\cite{DBLP:conf/icml/SonKKHY19}), 
 where a centralized learner trains distributed policies that are subsequently executed independently. 
 While facilitating decentralized runtime operation and avoiding single points of failure, 
 this approach typically entails an offline learning phase (using simulators and centralized computation) 
 followed by online execution with no further learning. 
 This inherent limitation hinders adaptation to changing environments or tasks at runtime.  
% 
While fully independent learners~\cite{abed2016comparison} offer a potential solution for online learning,
 they often underperform centralized counterparts due to instability in training and a lack of differentiation between agents and their interactions with the environment. 
%
However, in many real-world multi-agent systems, 
 agents only interact with a limited subset of the overall system (also called \emph{networked agents}),
 as exemplified by limited sensing ranges in swarm robotics or localized interactions.  
 This observation also holds true in other natural systems, such as animal herds.
%
Therefore, 
this paper proposes and formalizes a family of distributed learning methodologies based on neighborhood information.
This approach aims to: 
i) incorporate richer information compared to independent learners, and
ii) facilitate the emergence of collective behaviors through local interactions between networked agents.
Specifically, we investigate three distinct neighboring-based training strategies encompassing both neural network-based approaches 
(e.g., neighboring averaging and consensus) 
and experience-based methods inspired by centralized experience sharing~\cite{DBLP:conf/nips/ChristianosSA20}. 
We demonstrate the effectiveness of these strategies in comparison to both centralized and fully distributed approaches, 
highlighting how neighborhood-based policies offer a compelling compromise between these two extremes.

@@

These algorithms often exhibit improved scalability and communication efficiency compared to fully centralized methods, 
but still need lack in capturing the global structure of the environment and the interactions between agents.
\paragraph{Centralized Actor-Critic Methods}
\gianlu{
  todo -- discuss the main three solution, policy based (MAPPO, MADDPG, etc.), value based (QMIX, VDQN, etc.) and mean field (mean-q learning with field-informed RL).
}
\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning. 
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on deep Q-learning as the underlying learning algorithm,
and we consider the following strategies:
 k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.

>>>>

\sloppy
Many prominent approaches in this field employ a centralized training and decentralized execution paradigm
 (e.g.,MAPPO~\cite{yu2022surprising}, MADDPG~\cite{DBLP:conf/nips/LoweWTHAM17}, QMIX~\cite{DBLP:conf/icml/RashidSWFFW18}, QTRAN~\cite{DBLP:conf/icml/SonKKHY19}), 
 where a centralized learner trains distributed policies that are subsequently executed independently. 
 While facilitating decentralized runtime operation and avoiding single points of failure, 
 this approach typically entails an offline learning phase (using simulators and centralized computation) 
 followed by online execution with no further learning. 
 This inherent limitation hinders adaptation to changing environments or tasks at runtime.  
% 
While fully independent learners~\cite{abed2016comparison} offer a potential solution for online learning,
 they often underperform centralized counterparts due to instability in training and a lack of differentiation between agents and their interactions with the environment. 
%
However, in many real-world multi-agent systems, 
 agents only interact with a limited subset of the overall system (also called \emph{networked agents}),
 as exemplified by limited sensing ranges in swarm robotics or localized interactions.  
 This observation also holds true in other natural systems, such as animal herds.
%
Therefore, 
this paper proposes and formalizes a family of distributed learning methodologies based on neighborhood information.
This approach aims to: 
i) incorporate richer information compared to independent learners, and
ii) facilitate the emergence of collective behaviors through local interactions between networked agents.
Specifically, we investigate three distinct neighboring-based training strategies encompassing both neural network-based approaches 
(e.g., neighboring averaging and consensus) 
and experience-based methods inspired by centralized experience sharing~\cite{DBLP:conf/nips/ChristianosSA20}. 
We demonstrate the effectiveness of these strategies in comparison to both centralized and fully distributed approaches, 
highlighting how neighborhood-based policies offer a compelling compromise between these two extremes.

@@

These algorithms often exhibit improved scalability and communication efficiency compared to fully centralized methods, 
but still need lack in capturing the global structure of the environment and the interactions between agents.
\paragraph{Centralized Actor-Critic Methods}
\gianlu{
  todo -- discuss the main three solution, policy based (MAPPO, MADDPG, etc.), value based (QMIX, VDQN, etc.) and mean field (mean-q learning with field-informed RL).
}
\paragraph{Centralized Actor-Critic Methods} CTDE is a prominent paradigm in MARL~\cite{ning2024survey}, 
leveraging centralized information during training to learn decentralized policies for execution. 
This approach addresses the challenges of partial observability and non-stationarity while maintaining the scalability of decentralized execution. % 
Actor-critic methods are well-suited for the CTDE paradigm, 
where a centralized critic provides feedback to decentralized actors~\cite{DBLP:conf/icml/RashidSWFFW18,DBLP:journals/corr/SunehagLGCZJLSL17,yu2022surprising,DBLP:conf/nips/LoweWTHAM17}. 
% 
The most influential algorithms in this area are the one called policy-based methods (i.e., the ones that learn a policy directly).
%
Multi-agent deep deterministic policy Gradient (MADDPG)~\cite{DBLP:conf/nips/LoweWTHAM17} employs a centralized critic that has access to the states and actions of all agents, providing more informative feedback for policy updates. 
Each agent then learns a deterministic policy based on its local observations. % 
Building on the actor-critic framework, Multi-Agent Proximal Policy Optimization (MAPPO)~\cite{yu2022surprising} extends the centralized critic concept to the policy optimization setting. 
MAPPO leverages a centralized critic to estimate advantage functions for policy updates, while individual agents learn stochastic policies. 
This approach has shown promising results in various cooperative and competitive MARL tasks, exhibiting improved sample efficiency and robustness compared to MADDPG. % 
Another class of algorithms, called value-based methods, learn a value function that estimates the expected return for each agent.
%
Value Decomposition Networks (VDN)~\cite{DBLP:journals/corr/SunehagLGCZJLSL17} and QMIX~\cite{DBLP:conf/icml/RashidSWFFW18} are representative algorithms in this category. 
VDN factorizes the joint Q-function into a sum of individual agent Q-functions, simplifying the learning process. 
QMIX improves upon VDN by learning a monotonic mixing function that combines individual Q-values, 
allowing for more complex representations of joint action values. %

Another approach to address the scalability challenge in MARL is \emph{mean field reinforcement learning}, i.e., the one that models the interaction between agents through a mean field, simplifying the multi-agent problem into a single-agent problem against an average opponent.
%
In this context, mean field q-learning (Mean-Q)~\cite{DBLP:conf/icml/YangLLZZW18} has been proposed.
This approach scales well to large numbers of agents and has been successful in various applications. 
% 
There are also modern solution which leverages collective abstraction to reduce the complexity of the problem, such as the one proposed in field-informed reinforcement learning.
\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning. 
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on deep Q-learning as the underlying learning algorithm,
and we consider the following strategies:
 k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.
